---
title: "Channel selection"
author: "Dominik Klepl"
output: html_document
---
```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggthemes)
library(foreach)
library(doParallel)
library(magrittr)

#setup parallel processing
numCores = 12
registerDoParallel(numCores)

dir.create("channel_selection")
```

## Entropy selection
Select N channels with the highest entropy
```{r}
#compute Shannon entropy
shannon = function(x){
  dens = density(x)$y
  return(entropy::entropy(dens))
}

select_n_ent = function(folder){
  files = list.files(folder, full.names = T)
  
  shannon = function(x){
    dens = density(x)$y
    return(entropy::entropy(dens))
  }
  
  result = foreach(i=1:length(files), .combine = "rbind",
                   .packages = c("dplyr","tidyr")) %dopar% {
    load(files[i])
    data %>% 
      select(-Time) %>%
      summarise_all(shannon)
  }
  
  order = result %>%
    summarise_all(mean) %>%
    pivot_longer(cols = 1:23) %>%
    arrange(-value)
  return(order)
}

folders = list.files("data", full.names = T)[2:7]

ent_selections = list()
for(i in 1:length(folders)){
  selected  = select_n_ent(folders[i])
  name = strsplit(strsplit(folders[i], "/")[[1]][2], "_")[[1]][2]
  ent_selections[[name]] = selected
  print(name)
}

save(ent_selections, file = "channel_selection/entropy_based.RData")
stopImplicitCluster()
```

# SVM Selection
Select the most important channels based on permutation importance. More specifically, train SVM using all channels as features to predict diagnosis (AD/HC). Compute permutation importance of each channel and remove channel with lowest importance and re-train SVM. Continue while accuracy is improving. 
```{r}
library(caret)
library(foreach)
library(doParallel)
library(dplyr)
```

## Load and merge files (from one frequency band)
```{r}
data_folders = list.files("data", full.names = T)[2:7]

registerDoParallel(cores = 6)

dir.create("data/merged_clean")
for(folder in data_folders){
  files = list.files(folder, full.names = T)
  all = foreach (i = 1:length(files), .combine = "rbind") %dopar% {
    load(files[i])
    data = as.data.frame(eegkit::eegresample(data, (nrow(data)/2000)*250))
    data = data[1:250,-1]
    info = strsplit(strsplit(files[i],"/")[[1]][3],"_")[[1]]
    info[4] = gsub(".RData", "", info[4])
    
    data$ID = info[1]
    data$diagnosis = info[2]
    data$condition = info[3]
    data$epoch = info[4]
    data
  }
  save_as = paste0("data/merged_clean/",strsplit(folder,"_")[[1]][2], ".RData")
  save(all, file = save_as)
}
stopImplicitCluster()
```

## Train SVM
```{r}
select_features = function(file){
  load(file)
  
  all = all %>% select(-condition, -epoch) %>%
    mutate(diagnosis = as.factor(diagnosis))
  
  folds = groupKFold(all$ID, 5)
  all = select(all, -ID)
  
  fitControl = trainControl(method = "cv",number = 5, allowParallel = T, index = folds)
  
  #backward feature selection
  acc_best = 0
  acc_last = 0
  results = data.frame()
  data = all
  while((acc_best <= acc_last)==T){
    model = train(diagnosis ~ ., data = data, 
                   method = 'svmRadial',
                  preProcess = c("center", "scale"),
                   trControl = fitControl)
    importance = varImp(model)$importance
    importance = arrange(data.frame(var = rownames(importance), imp = importance$AD), imp)
    acc_last = max(model$results$Accuracy)
    remove = importance$var[1]
    importance = arrange(importance, -imp)
    result = cbind(paste(importance$var, sep = "", collapse = "+"), acc_last)
    data = dplyr::select(data, -remove)
    
    if(acc_best<acc_last){
      acc_best = acc_last
      results = rbind(results, result)
      print(acc_best)
    }
  }
  
  colnames(results) = c("Vars", "Accuracy")
  name = strsplit(strsplit(file, "/")[[1]][3],"\\.")[[1]][1]
  save_as = paste0("channel_selection/SVM/SVM_", name,".csv")
  readr::write_csv(results, save_as)
}
```

Loop through all frequency bands - i.e. make unique channel selection for each frequency band
```{r}
files = list.files("data/merged_clean", full.names = T)

registerDoParallel(numCores)
foreach(f = files) %do% {
  print(f)
  select_features(f)
}

stopImplicitCluster()
```

## Reformat the model channel selection
Combine the channel selections into one file
```{r}
files = list.files("channel_selection/SVM", full.names = T)

SVM_selections = foreach(i = 1:length(files), .combine = "cbind") %do% {
  d = read.csv(files[i])
  selected = d$Vars[nrow(d)]
  acc = round(d$Accuracy[nrow(d)],2)
  selected = strsplit(selected, "\\+")[[1]]
  
  n_removed = 23 - length(selected)
  add_NA = rep(NA, n_removed)
  c(acc, selected, add_NA)
} %>% as.data.frame()
colnames(SVM_selections) = c("alpha", "beta", "delta", "full", "gamma", "theta")


#reformat to rows - each selection is one row
SVM_selections = as.data.frame(t(SVM_selections))
SVM_selections$band = rownames(SVM_selections)
colnames(SVM_selections)[1] = "accuracy"

readr::write_csv(SVM_selections, "channel_selection/SVM_based.csv")
```

#Find intersection between the methods
Produce final channel selections by finding 10 channels selected by both methods (entropy and SVM). If there's less than 10 common channels, the SVM-selection is prefered.
```{r}
library(dplyr)
load("channel_selection/entropy_based.RData")
entropy = foreach(i=names(ent_selections), .combine = "rbind")%do%{
  ent_selections[[i]]$name %>% as.data.frame() %>% t() %>% as.data.frame() %>%
    mutate(band = i)
}

SVM = read.csv("channel_selection/SVM_based.csv") %>% select(-accuracy)

bands = c("alpha", "beta", "delta", "full", "gamma", "theta")


overlap = foreach(b = bands) %do% {
  ent = entropy %>% dplyr::filter(band==b) %>% select(-band) %>% as.character() %>% .[1:15]
  SVM_one = SVM %>% dplyr::filter(band==b) %>% select(-band) %>% as.character() %>% na.omit() %>% .[1:15]
  intersect(SVM_one, ent)
}

names(overlap) = bands

#final selection
final = foreach(b=names(overlap), .combine = "rbind")%do%{
  i = overlap[[b]]
  if(length(i)==10){
    res = c(i,b)
  }
  if(length(i)>10){
    res = c(i[1:10],b)
  }
  if(length(i)<10){
    N = 10-length(i) #how many to add from SVM selection
    SVM_add = SVM %>% filter(band==b) %>% select(-band) %>% as.character()
    add = setdiff(SVM_add,i)[1:N]
    res = c(i,add,b)
  }
  res
} %>% as.data.frame()

readr::write_csv(final, "channel_selection/final_selection.csv")
```

## Plot final selection on montage
Creates Figure 2.
```{r}
source("scripts/preprocess.R")
library(dplyr)
data = readr::read_csv("channel_selection/final_selection.csv")

locs = get_channel_loc()
```

Create montage
```{r}
names = c("F8_F4","F7_F3","F4_C4","F3_C3","F4_FZ","F3_FZ","FZ_CZ","T4_C4","T3_C3","C4_CZ","C3_CZ","CZ_PZ","C4_P4","C3_P3","T4_T6","T3_T5","P4_PZ","P3_PZ","T6_O2","T5_O1","P4_O2","P3_O1","O1_O2")

#get positions of individual channels
data(eegcoord)
pair_locs = matrix(NA, length(names)*2, 3)

single_channels = sapply(names, strsplit, split="_") %>% unlist() %>% unique()
names(single_channels) = NULL

single_df = eegcoord[single_channels, 4:5]
colnames(single_df) = c("x","y")
single_df$channel = single_channels

single_df[single_df$channel=="T3",1:2] = eegcoord["T7",4:5]
single_df[single_df$channel=="T4",1:2] = eegcoord["T8",4:5]
single_df[single_df$channel=="T5",1:2] = eegcoord["P7",4:5]
single_df[single_df$channel=="T6",1:2] = eegcoord["P8",4:5]

geom_head()+
  geom_text(data=single_df, aes(x,y, label=channel))+
  geom_point(data=single_df, aes(x,y), shape=1, size=9)

#construct lines between channel pairs
channel_loc = matrix(0, length(names),5)
colnames(channel_loc) = c("x_start", "y_start", "x_end","y_end", "channel")
channel_loc[,"channel"] = names

for(i in 1:length(names)){
  pair_name = strsplit(names[i],"_")[[1]]
  
  channel_loc[i,1:2] = as.numeric(single_df[single_df$channel==pair_name[1],1:2])
  channel_loc[i,3:4] = as.numeric(single_df[single_df$channel==pair_name[2],1:2])
}

channel_loc = as.data.frame(channel_loc)

channel_loc$x_start = as.numeric(channel_loc$x_start)
channel_loc$y_start = as.numeric(channel_loc$y_start)
channel_loc$x_end = as.numeric(channel_loc$x_end)
channel_loc$y_end = as.numeric(channel_loc$y_end)

channel_loc$channel = gsub("_", "", channel_loc$channel)

montage = geom_head()+
  geom_point(data=single_df, aes(x,y), size=16, fill="grey", shape=21,alpha=0.3)+
  geom_text(data=single_df, aes(x,y, label=channel), size=7.5)
```

Plot selections
```{r}
plot_selection = function(selected, title) {
  channel_loc$selected = ifelse(channel_loc$channel %in% selected, "selected", "1")
  
  channel_plot = montage+
    geom_segment(data=channel_loc, aes(x = x_start, y = y_start, xend=x_end, yend=y_end, colour=selected), alpha=0.3, size=2.5, lineend = "butt", linejoin = "bevel")+
    scale_color_manual(breaks = c("Selected"),
                       values = c("grey", "blue"))+
    theme(plot.title = element_text(hjust = 0.5, size=25))+
    labs(title = title)
  return(channel_plot)
}
```

```{r}
selections = read.csv("channel_selection/final_selection.csv")
colnames(selections)[11] = "band"

bands = unique(selections$band)

library(foreach)
final_plots = foreach(b = bands) %do% {
  selection = selections %>% dplyr::filter(band == b) %>% select(-band) %>% as.character() %>% na.omit()
  plot_selection(selection, b)
}
names(final_plots) = bands

#put the different selections into a single figure
library(patchwork)
balanced_plot = (final_plots$alpha | final_plots$beta | final_plots$delta)/
(final_plots$gamma | final_plots$theta | final_plots$full)
balanced_plot

ggsave("figures/selected_channels.png", balanced_plot, width = 9.5, height = 7)
ggsave("figures/pdf/selected_channels.pdf", balanced_plot, width = 9.5, height = 7)
```