---
title: "Energy Landscape: Pipeline"
author: "Dominik Klepl"
output:
  pdf_document: default
  html_document: default
---

First we define a large function that fits pMEM, uses J parameters of pMEM to train SVM classifier (connectivity-based), extracts energy values of all states and trains SVM classifier (energy-based) and extract energy landscape features.
This function is then used to optimise sampling frequency and window size so that the predictive accuracy is maximised.

#Large pipeline function
```{r}
run_all = function(name="a", window=1, down=5, base="results/",
               k = 10, repeats = 100){
  #name - name of directory to save results
  #window - window size to be used for binarisation of signals (in seconds), the window size is then computed as window*down
  #down - sampling frequency to which to downsample
  #base - root directory to save results
  #k - for repeated k-fold cross-validation
  #repeats - number of repetitions for repeated k-fold cross-validation
  
  
  source("scripts/pMEM.R")
  source("scripts/ELA.R")

  #create folders for saving the results
  root = paste0(base, name)
  fold_pMEM = paste0(root, "/pMEM")
  fold_pMEM_feat = paste0(fold_pMEM, "_features")
  dir.create(base)
  dir.create(root)
  dir.create(fold_pMEM)
  dir.create(fold_pMEM_feat)
  
  #downsample EEG to desired sampling frequency
  downsample = function(x, freq){
    len = length(x)/2000
    
    n = floor(len*freq)
    return(eegkit::eegresample(x, n))
  }
  
  #estimate pMEM for one frequency band
  pMEM_band = function(wave_band, selection, size, down){
    input = paste0("data/clean_", wave_band) #change to location of your data
    files = list.files(input, full.names = T)
    
    pMEM_results = matrix(NA, length(files), 7)
    colnames(pMEM_results) = c("ID", "diagnosis", "condition", "epoch", "band", "r", "R2") #29
    for(i in 1:length(files)){
      load(files[i])
      data = data[,selection]
      
      #downsample
      data = apply(data, 2, downsample, freq=down) %>% as.data.frame()
      
      results = run_pMEM(data, size)
      
      info = strsplit(strsplit(files[i],"/")[[1]][3],"_")[[1]]
      ID = info[1]
      diagnosis = info[2]
      condition = info[3]
      epoch = strsplit(info[4],"\\.")[[1]][1]
      
      #save to data/pMEM_3_500Hz
      filename = paste0(wave_band,"_",strsplit(files[i],"/")[[1]][3])
      save_to = paste0(fold_pMEM,"/",filename)
      save(results, file = save_to)
      
      results_row = cbind(ID, diagnosis, condition, epoch, wave_band, results$metrics$r, results$metrics$R2)
      
      pMEM_results[i,] = results_row
    }
    
    pMEM_results = as.data.frame(pMEM_results)
    pMEM_results = na.omit(pMEM_results)
    
    return(pMEM_results)
  }
  
  
  #load the 10 selected channels
  selections = readr::read_csv("channel_selection/final_selection.csv")
  bands = c("alpha", "beta", "delta", "full", "gamma", "theta")
  
  #estimate pMEM for each frequency band
  registerDoParallel(6)
  result = foreach (b = bands, .combine = "rbind", .packages = "tidyverse") %dopar% {
    source("scripts/pMEM.R")
    selection = selections %>% dplyr::filter(V11 == b)
    selection = as.character(selection[1,1:10]) %>% na.omit()
    r=pMEM_band(wave_band = b, selection = selection, size=window, down = down)
  }
  stopImplicitCluster()
  readr::write_csv(result, paste0(root,"/pMEM_metrics.csv"))
  
  #collect J and h parameters of pMEM to be used for ML
  files = list.files(fold_pMEM, full.names = T)
  bands = c("alpha", "beta", "delta", "full", "gamma", "theta")
  
  registerDoParallel(6)
  for(b in bands){
    Jh_features = foreach(f=files[grep(b,files)], .combine = "rbind", .packages = "tidyverse") %dopar% {
      load(f)
      J = results$parameters$J
      J[lower.tri(J)] = 0
      diag(J) = results$parameters$h + 1e-50
      J = reshape2::melt(J) %>% filter(value != 0) %>% unite("name",Var1:Var2)
      res = J$value
      names(res) = J$name
      res = as.data.frame(t(res))
      
      res$band = strsplit(f, "/")[[1]][4] %>% strsplit("_") %>% .[[1]] %>% .[1]
      res$ID = strsplit(f, "/")[[1]][4] %>% strsplit("_") %>% .[[1]] %>% .[2]
      res$diagnosis = strsplit(f, "/")[[1]][4] %>% strsplit("_") %>% .[[1]] %>% .[3]
      res$condition = strsplit(f, "/")[[1]][4] %>% strsplit("_") %>% .[[1]] %>% .[4]
      res$epoch = strsplit(f, "/")[[1]][4] %>% strsplit("_") %>% .[[1]] %>% .[5] %>% strsplit("\\.") %>% .[[1]] %>% .[1]
      res
    }
    
    save_to = paste0(fold_pMEM_feat,"/", b, ".csv")
    readr::write_csv(Jh_features, save_to)
  }
  stopImplicitCluster()
  
  #ML on connectivity
  library(caret)
  files = list.files(fold_pMEM_feat, full.names = T)
  
  fit_ctrl = trainControl(method = "repeatedcv",
                        number = k,
                        repeats = repeats,
                        verboseIter = F,
                        classProbs = T,
                        summaryFunction = twoClassSummary,
                        savePredictions = "final",
                        allowParallel = T)

  registerDoParallel(11)

  method = 'svmRadial'
  results = foreach(f = files, .combine = "rbind") %do% {
    df = read.csv(f)
    band = unique(df$band)
    
    EC = df %>% filter(condition=="EC") %>% select(-band, -condition, -epoch, -ID)
    EO = df %>% filter(condition=="EO") %>% select(-band, -condition, -epoch, -ID)
    
    model = train(diagnosis ~ .,
                      data = EC,
                      trControl = fit_ctrl,
                      method = method,
                  preProcess = c("scale", "center", "pca"),
                  metric = "ROC")
    resamp_EC = data.frame(band=band, 
                           condition = "EC", 
                           AUC = model$resample$ROC,
                           Sensitivity = model$resample$Sens,
                           Specificity = model$resample$Spec)
    
    model = train(diagnosis ~ .,
                      data = EO,
                      trControl = fit_ctrl,
                      method = method,
                  preProcess = c("scale", "center", "pca"),
                  metric = c("ROC"))
    resamp_EO = data.frame(band=band, 
                           condition = "EO", 
                           AUC = model$resample$ROC,
                           Sensitivity = model$resample$Sens,
                           Specificity = model$resample$Spec)
    
    result = rbind(resamp_EC, resamp_EO)
    result
  }
  stopImplicitCluster()
  readr::write_csv(results, paste0(root,"/ML_connectivity.csv"))
  
  #ML on energy
  
  ## collect energies of all states
  source("scripts/collect_energy_features.R")

  data = collect_energy(fold_pMEM, cores = 2)
  save_to = paste0(root,"/states_energy.RData")
  save(data, file = save_to)
  
  #models
  data$diagnosis = as.factor(data$diagnosis)
  bands = unique(data$band)
  
  fit_ctrl = trainControl(method = "repeatedcv",
                          number = k,
                          repeats = repeats,
                          verboseIter = F,
                          classProbs = T,
                          summaryFunction = twoClassSummary,
                          savePredictions = "final",
                          allowParallel = T)
  
  registerDoParallel(12)
  
  method = 'svmRadial'
  results = foreach(b = bands, .combine = "rbind") %do% {
    df = data %>% filter(band==b)
    EC = df %>% filter(condition=="EC") %>% select(-band, -condition, -epoch, -ID)
    EO = df %>% filter(condition=="EO") %>% select(-band, -condition, -epoch, -ID)
    
    model = train(diagnosis ~ .,
                      data = EC,
                      trControl = fit_ctrl,
                      method = method,
                  preProcess = c("scale", "center", "pca"),
                  metric = "ROC")
    resamp_EC = data.frame(band=b, 
                           condition = "EC", 
                           AUC = model$resample$ROC,
                           Sensitivity = model$resample$Sens,
                           Specificity = model$resample$Spec)
    
    model = train(diagnosis ~ .,
                      data = EO,
                      trControl = fit_ctrl,
                      method = method,
                  preProcess = c("scale", "center", "pca"),
                  metric = c("ROC"))
    resamp_EO = data.frame(band=b, 
                           condition = "EO", 
                           AUC = model$resample$ROC,
                           Sensitivity = model$resample$Sens,
                           Specificity = model$resample$Spec)
    result = rbind(resamp_EC, resamp_EO)
    result
  }
  stopImplicitCluster()
  readr::write_csv(results, paste0(root,"/ML_energy.csv"))
  
  #extract energy features
  source("scripts/ELA.R")
  library(foreach)
  library(doParallel)
  
  #construct energy landscape and extract features
  analyze_energy_landscape = function(file){
    source("scripts/ELA.R")
    load(file)
    
    fname = strsplit(file, "/")[[1]][4]
    band = strsplit(fname, "_")[[1]][1]
    ID = strsplit(fname, "_")[[1]][2]
    diagnosis = strsplit(fname, "_")[[1]][3]
    condition = strsplit(fname, "_")[[1]][4]
    epoch = strsplit(strsplit(fname, "_")[[1]][5], "\\.")[[1]][1]
    
    info = data.frame(band, ID, diagnosis, condition, epoch)
    
    
    #data.frame for storing results
    features = data.frame(minima = NA, E_diff = NA, basin_sd = NA, duration = NA)
    
    states_energy = get_state_energy(results$parameters)
    adj = get_adjacency(states_energy$state)
    
    #local minima related features
    minima = find_minima(adj, states_energy)
    features$minima = nrow(minima)
    
    features$E_diff = mean_energy_difference(minima)
    if(is.na(features$E_diff)){features$E_diff=0}
    
    global_min = minima$state[minima$energy==min(minima$energy)]
    
    #basin size
    basin_df = basin_size_estimator(adj, states_energy)
    #basin_plot = plot_basin(basin_df)
    features$basin_sd = basin_df$membership %>% 
      group_by(basin) %>%
      summarise(count = n()) %>%
      pull() %>%
      sd()
    
    if(is.na(features$basin_sd)){features$basin_sd= nrow(basin_df$membership)}
    
    #simulate transitions
    duration = NULL
    while(is_empty(duration)){
      x = simulate_transitions(2e4, adj, states_energy)
      duration = data.frame(state = x) %>%
        left_join(basin_df$membership, by="state") %>%
        group_by(basin) %>%
        summarise(duration = n()) %>%
        mutate(duration = duration/500)
      duration = duration$duration[duration$basin==global_min]
    }
    features$duration = duration
    
    features = cbind(info, features)
    return(features)
  }
  
  #loop through all files (within the folder with fitted pMEM)
  files = list.files(fold_pMEM, full.names = T)
  
  registerDoParallel(10)
  results = foreach(f = files, .combine = "rbind", .packages = c("tidyverse")) %dopar% {
    feat = try(analyze_energy_landscape(f))
    
    if(is.character(feat)){
      rep(NA,(9))
    } else {feat}
  }
  stopImplicitCluster()
  
  nas = which(is.na(results$band))
  
  for (n in nas){
    results[n,] = try(analyze_energy_landscape(files[n]))
  }
  
  readr::write_csv(results, paste0(root,"/energy_features.csv"))
}
```

#Try different sampling frequencies and window sizes
```{r}
sampling = c(500, 1000, 1500, 2000)
time = c(0.1,0.2,0.3,0.4,0.5,1,1.5,2,2.5,3,3.5,4)
source("scripts/full_analysis.R")

for(s in sampling){
  for(t in time){
    window = t*s
    name = paste0(window, "_",s,"Hz")
    if(dir.exists(paste0("results/",name))){
      print(name)
      next
    }else{
    run(name, window, s, base = "results/")
    print(name)
    }
  }
}
```

#Select the sampling frequency and window size
```{r}
TODO
```









