---
title: "Results"
author: "Dominik Klepl"
output: html_document
---


Analyse pMEM metrics - what is the average goodness-of-fit and does it differ across frequency bands,conditions and diagnosis.
```{r}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(ggpubr)
library(rstatix)

df = read.csv("time_window_results/5250_1500Hz/pMEM_metrics.csv")
df$diagnosis = as.factor(df$diagnosis)

hist(df$r)

df %>% anova_test(r ~ diagnosis+band*condition, wid=ID) %>% as_tibble() %>% mutate(p = round(p,3),
                                                                                   `F` = round(`F`,3))
mean(df$r)
sd(df$r)

m1 = lm(r ~ diagnosis*band*condition, df)
hist(m1$residuals)
```

# Connectivity vs energy models
Compare the models trained on connectivity features and energy values. We predicted that energy-based models are better.
```{r}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(rstatix)
library(ggpubr)
library(coin)
library(foreach)
root.dir = "results/5250_1500Hz/"

connect = read.csv(paste0(root.dir,"ML_connectivity.csv")) %>% mutate(type = "Connectivity")
energy = read.csv(paste0(root.dir,"ML_energy.csv")) %>% mutate(type = "Energy")
all_samples = rbind(connect, energy)
all_samples$band = factor(all_samples$band, levels = c("delta", "theta", "alpha", "beta", "gamma", "full"))
all_samples$type = as.factor(all_samples$type)
all_samples$condition = as.factor(all_samples$condition)

m = lm(AUC~type*band*condition, all_samples)
qqnorm(m$residuals)

all_samples %>% 
  anova_test(AUC ~ type*band*condition) %>% 
  as_tibble() %>%
  mutate(report = paste0("F(", DFn, ",", DFd,")=",round(`F`,3),",p=",round(p,3)))

report = all_samples %>%
  group_by(band, condition) %>%
  anova_test(AUC ~ type, wid = c(condition, band))%>% 
  adjust_pvalue() %>%
  add_significance() %>%
  as_tibble() %>%
  mutate(report = paste0("F(", DFn, ",", DFd,")=",round(`F`,3),",p=",round(p.adj,3)))

#interaction
ML_stats = all_samples %>%
  group_by(band, condition) %>%
  tukey_hsd(AUC ~ type) %>%
  add_xy_position(x="band", dodge = 0.8) %>%
  mutate(.y. = "AUC",
         p.adj.signif = ifelse(p.adj<=0.05,"*","ns"))

plot = ggbarplot(all_samples, x="band", y="AUC",fill = "type", position = position_dodge(0.8), add="mean_ci", facet.by = "condition", add.params = list(width=0.35))+
  theme_few()+
  scale_fill_manual(values = c("#BA0C2F","#00629B"), name = "Predicted by")+
  labs(x = "Frequency band", y = "AUC")+
  scale_x_discrete(labels=c("alpha" = expression(alpha), "beta" = expression(beta),
                              "delta" = expression(delta), "full" = "f", 
                            "gamma" = expression(gamma),
                            "theta" = expression(theta)))+
  coord_cartesian(ylim=c(0.45,0.98))+
  theme(axis.title = element_text(size=20),
        axis.text.x = element_text(size=22),
        axis.text.y = element_text(size=20),
        strip.text = element_text(size=20),
        legend.title = element_text(size=18),
        legend.text = element_text(size=17),
        legend.position = "bottom")
plot

ggsave("figures/ML_connectivity_vs_energy.png", plot, width = 8, height = 4)
ggsave("figures/pdf/ML_connectivity_vs_energy.pdf", plot, width = 8, height = 4)

#best energy performance
energy %>% filter(band=="full") %>%
  summarise(auc = mean(AUC),
            sens = mean(Sensitivity),
            spec = mean(Specificity))
```


# Energy landscape features

4 features of energy landscape were extracted: number of local minima, energy difference between global minimum and local minima, SD of basin size and time in basin of global minimum.
ANOVA is used to test for differences between diagnosis, conditions and frequency bands (+their ineraction). Significant effects are analysed post-hoc.
```{r}
library(ggplot2)
library(ggthemes)
library(patchwork)
library(ggpubr)
library(rstatix)
library(dplyr)

df = read.csv("time_window_results/5250_1500Hz/energy_features.csv")
df$diagnosis = as.factor(df$diagnosis)
df$condition = as.factor(df$condition)
df$ID = as.factor(df$ID)
df$band = factor(df$band, levels = c("delta", "theta", "alpha", "beta", "gamma", "full"))
```

## Local minima
```{r}
m1 = lm(minima ~ diagnosis*band*condition, df)
qqnorm(m1$residuals)

df %>% 
  anova_test(minima ~ diagnosis*band*condition, wid = ID) %>% 
  as_tibble() %>%
  mutate(report = paste0("F(", DFn, ",", DFd,")=",round(`F`,3),",p=",round(p,3)))

#if significant interaction
#stats = df %>% group_by(band, condition) %>% tukey_hsd(minima ~ diagnosis)

minima = ggbarplot(df, "band", "minima", fill = "diagnosis", facet.by = "condition",
          add="mean_se", add.params = list(width=0.35),
          position = position_dodge(0.8))+
    theme_few()+
    scale_fill_manual(values = c("#BA0C2F","#00629B"), name = "Diagnosis")+
  labs(x = NULL,y = "Local Minima")+
  scale_x_discrete(labels=c("alpha" = expression(alpha), 
                            "beta" = expression(beta),
                            "delta" = expression(delta), 
                            "full" = "f", 
                            "gamma" = expression(gamma),
                            "theta" = expression(theta)))+
  coord_cartesian(ylim=c(5,10.1))+
  theme(axis.title = element_text(size=18),
        axis.text = element_text(size=18),
        strip.text.x = element_text(size=18),
        legend.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.position = "bottom")
minima
```

Energy difference
```{r}
m1 = lm(E_diff ~ diagnosis*band*condition, df)
qqnorm(m1$residuals)

df %>% 
  anova_test(E_diff ~ diagnosis*band*condition, wid = ID) %>% 
  as_tibble() %>%
  mutate(report = paste0("F(", DFn, ",", DFd,")=",round(`F`,3),",p=",round(p,3)))

#if significant interaction
stats = df %>% group_by(band) %>% anova_test(E_diff ~ diagnosis)

E_diff = ggbarplot(df, "band", "E_diff", fill = "diagnosis", facet.by = "condition",
          add="mean_se", add.params = list(width=0.35),
          position = position_dodge(0.8))+
    theme_few()+
    scale_fill_manual(values = c("#BA0C2F","#00629B"), name = "Diagnosis")+
  labs(x = NULL,y = "Energy Difference")+
  scale_x_discrete(labels=c("alpha" = expression(alpha), "beta" = expression(beta),
                              "delta" = expression(delta), "full" = "f", 
                            "gamma" = expression(gamma),
                            "theta" = expression(theta)))+
  coord_cartesian(ylim = c(0.9,1.65))+
  theme(axis.title = element_text(size=18),
        axis.text = element_text(size=18),
        strip.text.x = element_text(size=18),
        legend.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.position = "bottom")
E_diff

report = stats  %>% 
  as_tibble() %>%
  mutate(report = paste0("F(", DFn, ",", DFd,")=",round(`F`,3),",p=",round(p,3)))
```

## Basin sd
```{r}
#make it normal
hist(df$basin_sd)
df$basin_sd = sqrt(df$basin_sd)

m1 = lm(basin_sd ~ diagnosis*band*condition, df)
qqnorm(m1$residuals)

df %>% 
  anova_test(basin_sd ~ diagnosis*band*condition, wid = ID) %>% 
  as_tibble() %>%
  mutate(report = paste0("$(F(", DFn, ",", DFd,")=",round(`F`,3),",p=",round(p,3),")$"))

#band:condition
df %>% group_by(band) %>% anova_test(basin_sd ~ condition) %>%as_tibble() %>%
  mutate(report = paste0("$(F(", DFn, ",", DFd,")=",round(`F`,3),",p=",round(p,3),")$"))
df %>% group_by(band) %>% tukey_hsd(basin_sd ~ condition)


#group:band
stats = df %>% group_by(band) %>% anova_test(basin_sd ~ diagnosis)

#band:condition
#stats = df %>% group_by(band) %>% tukey_hsd(basin_sd ~ condition)

basin_sd = ggbarplot(df, "band", "basin_sd", fill = "diagnosis", facet.by = "condition",
          add="mean_se", add.params = list(width=0.35),
          position = position_dodge(0.8))+
    theme_few()+
    scale_fill_manual(values = c("#BA0C2F","#00629B"), name = "Diagnosis")+
  labs(x = NULL,y = "sd(Basin size)")+
  scale_x_discrete(labels=c("alpha" = expression(alpha), "beta" = expression(beta),
                              "delta" = expression(delta), "full" = "f", 
                            "gamma" = expression(gamma),
                            "theta" = expression(theta)))+
  coord_cartesian(ylim = c(9.4,13.85))+
  theme(axis.title = element_text(size=18),
        axis.text = element_text(size=18),
        strip.text.x = element_text(size=18),
        legend.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.position = "bottom")
basin_sd

report = stats  %>% 
  as_tibble() %>%
  mutate(report = paste0("$(F(", DFn, ",", DFd,")=",round(`F`,3),",p=",round(p,3),")$"))
```

## Duration
```{r}
hist(sqrt(df$duration))
hist(df$duration)

df$duration = sqrt(df$duration)
m1 = lm(duration ~ diagnosis*band*condition, df)
qqnorm(m1$residuals)

df %>% 
  anova_test(sqrt(duration) ~ diagnosis*band*condition, wid = ID) %>% 
  as_tibble() %>%
  mutate(report = paste0("$(F(", DFn, ",", DFd,")=",round(`F`,3),",p=",round(p,3),")$"))

#band:condition
df %>% group_by(band) %>% anova_test(duration ~ condition) %>% as_tibble() %>%
  mutate(report = paste0("$(F(", DFn, ",", DFd,")=",round(`F`,3),",p=",round(p,3),")$"))
ggbarplot(df, "band", "duration", fill = "condition",
          add="mean_se", add.params = list(width=0.35),
          position = position_dodge(0.8))+
    theme_few()+
    scale_fill_manual(values = c("#BA0C2F","#00629B"), name = "Diagnosis")+
  labs(x = NULL,y = "Time in Global Minimum")+
  scale_x_discrete(labels=c("alpha" = expression(alpha), "beta" = expression(beta),
                              "delta" = expression(delta), "full" = "f", 
                            "gamma" = expression(gamma),
                            "theta" = expression(theta)))

#group:band
stats = df %>% group_by(band) %>% anova_test(duration ~ diagnosis)

#band:condition
df %>% group_by(band) %>% tukey_hsd(duration ~ condition)

duration = ggbarplot(df, "band", "duration", fill = "diagnosis", facet.by = "condition",
          add="mean_se", add.params = list(width=0.35),
          position = position_dodge(0.8))+
    theme_few()+
    scale_fill_manual(values = c("#BA0C2F","#00629B"), name = "Diagnosis")+
  labs(x = NULL,y = "Time in GM")+
  scale_x_discrete(labels=c("alpha" = expression(alpha), "beta" = expression(beta),
                              "delta" = expression(delta), "full" = "f", 
                            "gamma" = expression(gamma),
                            "theta" = expression(theta)))+
  coord_cartesian(ylim = c(3.25,4.63))+
  theme(axis.title = element_text(size=18),
        axis.text = element_text(size=18),
        strip.text.x = element_text(size=18),
        legend.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.position = "bottom")
duration

report = stats  %>% 
  as_tibble() %>%
  mutate(report = paste0("$(F(", DFn, ",", DFd,")=",round(`F`,3),",p=",round(p,3),")$"))
```

Put the four plots together into a large figure (as in Figure 5)
```{r}
library(patchwork)
final_plot = minima + E_diff+ basin_sd + duration + 
  plot_layout(guides = 'collect') + plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(size=24),
        legend.position = "bottom",
        axis.text = element_text(size=22),
        axis.title = element_text(size=20))

ggsave("final_results/linear_differences.pdf", final_plot, width = 12, height = 7, dpi = 300)
ggsave("final_results/Figure5.pdf", final_plot, width = 12, height = 7, dpi = 300)
```